FATE summary results

### 2019-12-11 paper draft.Rmd
Typo
Updated plots for ms. revisions

```{r, echo=F}
# trends
plot.trend <- data.frame(year=1972:2017, trend=as.vector(rotated$trends_mean),
                         lo=as.vector(rotated$trends_lower),
                         hi=as.vector(rotated$trends_upper))

plot.b <- ggplot(plot.trend, aes(year, trend)) +
  theme_bw() +
  geom_line(color=cb[3]) +
  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
  geom_hline(yintercept = 0) + ylab("Trend value") +
  theme(axis.title.x = element_blank()) + 
  ggtitle("b) Trend")

png("combined GOA climate loading and trend plot.png", 8, 4, units = "in", res=300)

@@

clim.regimes <- data.frame(year=1950:2019, probability=kk$data$median)

@@

biol.regimes <- data.frame(year=1972:2017, probability=kk$data$median)

biol.regimes$data.set <- "Biology 1972-2017"

@@

@@

```{r, echo=F}
GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
GOA.ich$summary
```

@@

# reload the version of data used
# load updated ewidata!
# load updated ewidata!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))

# limit to GOA
GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
unique(GOA$code)

# limit to ichthy

keep <- grep("ICH", GOA$code)

GOA <- GOA[keep,]

@@

rotated = rotate_trends(GOA.ich$best_model) 
new.names <- c("sandlance", "ronquils", "walleye.pollock", "Pacific.cod", "flathead.sole", "n.rock.sole", "rockfish")

@@

@@

ichthyo.regimes <- data.frame(year=plot.trend$year, probability=kk$data$median)

ichthyo.regimes$data.set <- "Ichthyoplankton 1980-2015"

plot.regimes <- rbind(clim.regimes, biol.regimes, ichthyo.regimes)

>>>>

```{r, echo=F}
# trends
# plot.trend <- data.frame(year=1972:2017, trend=as.vector(rotated$trends_mean),
#                          lo=as.vector(rotated$trends_lower),
#                          hi=as.vector(rotated$trends_upper))
# 
# plot.b <- ggplot(plot.trend, aes(year, trend)) +
#   theme_bw() +
#   geom_line(color=cb[3]) +
#   geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
#   geom_hline(yintercept = 0) + ylab("Trend value") +
#   theme(axis.title.x = element_blank()) + 
#   ggtitle("b) Trend")
# 
# png("combined GOA climate loading and trend plot.png", 8, 4, units = "in", res=300)
# 
# ggpubr::ggarrange(plot.a, plot.b, ncol=2)
# 
# dev.off()
```

@@

clim.regimes <- data.frame(year=1950:2019, probability=kk$data$median, lwr=kk$data$lwr, upr=kk$data$upr)

@@

biol.regimes <- data.frame(year=1972:2019, probability=kk$data$median, lwr=kk$data$lwr, upr=kk$data$upr)

biol.regimes$data.set <- "Biology 1972-2019"

@@

We're interested in testing the hypothesis that the slope on sst was greater during 1972:1988 (high Aleutian Low variance) than during 1989:2013 (low Aleutian Low variance). Also interested in asking which of these periods the 2014-2016 heat wave years are most similar to.

@@

```{r, echo=F}
# GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
GOA.ich <- readRDS("/Users/MikeLitzow/Documents/github/ewi-in-recovery/ewi-in-recovery/all_plankton_11.20.19.one.trend.rds") 
GOA.ich$summary
```

@@

# reload the version of data used
# load updated ewidata!
# load updated ewidata!
# ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
# 
# # limit to GOA
# GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
# unique(GOA$code)
# 
# # limit to ichthy
# 
# keep <- grep("ICH", GOA$code)
# 
# GOA <- GOA[keep,]

# recover the years and names for plots!
# using GOA from 'updated plankton model.R'
sub_data = GOA

@@

rotated = rotate_trends(GOA.ich$best_model) 
# new.names <- c("sandlance", "ronquils", "walleye.pollock", "Pacific.cod", "flathead.sole", "n.rock.sole", "rockfish")
# 
#  check <- data.frame(names=names, newnames=new.names)
#  check

@@

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(11)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 5)

@@

ichthyo.regimes <- data.frame(year=plot.trend$year, probability=kk$data$median, lwr=kk$data$lwr, upr=kk$data$upr)

ichthyo.regimes$data.set <- "Plankton 1981-2019"

plot.regimes <- rbind(clim.regimes, biol.regimes, ichthyo.regimes)
plot.regimes$jitter <- ifelse(plot.regimes$data.set=="Biology 1972-2019" & plot.regimes$probability>0.99, 1.002, plot.regimes$probability)

plot.regimes$order <- ifelse(plot.regimes$data.set=="Plankton 1981-2019", 1,
                             ifelse(plot.regimes$data.set=="Biology 1972-2019", 2, 3))

plot.regimes$data.set <- reorder(plot.regimes$data.set, plot.regimes$order)

ggplot(plot.regimes, aes(year, jitter, color=data.set, fill=data.set)) +
  theme_bw() +
  geom_line() +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.2, color=NA) +
  scale_color_manual(values=cb[c(2,7,6)], guide=guide_legend(reverse=T)) + scale_fill_manual(values=cb[c(2,7,6)], guide=guide_legend(reverse=T)) +
  xlab("Year") + ylab("Probability") +
  theme(legend.title = element_blank(), legend.position = 'top')

ggsave("updated HMM regime plot.png", width = 6, height = 4, units='in)
```
